---
title: "Testing some stereotypes"
format: pdf
Author: Tuba
---

## 2.3: Testing some stereotypes

### Loading the data

```{r}
# Import the dataset
Snore <- read.delim("snore.txt", header = TRUE)

head(Snore)
str(Snore) # structure of the dataset
summary(Snore)
```

```{r}
# Check the categories of categorical variables
unique(Snore$sex)
unique(Snore$snore)
unique(Snore$tabac)
```

```{r}
# Count observations in each category
table(Snore$sex)
table(Snore$snore)
table(Snore$tabac)
```
### First stereotype: Are snorers fatter?

```{r}
### Are snorers fatter?

# Compare weight between snorers and non-snorers
boxplot(weight ~ snore,
        data = Snore,
        xlab = "Snoring category",
        ylab = "Weight",
        main = "Weight comparison between snorers and non-snorers")
```

```{r}
# Calculate median weight in each group
tapply(Snore$weight, Snore$snore, median)
```
```{r}
# Statistical analysis
# Check normality of weight in each snoring group
shapiro.test(Snore$weight[Snore$snore == "N"])
shapiro.test(Snore$weight[Snore$snore == "O"])

# Compare weight between the two independent groups
wilcox.test(weight ~ snore,
            data = Snore,
            exact = FALSE)
```

**Result – Weight comparison** 
The median weight was 94 among non-snorers and 96 among snorers. The Wilcoxon rank-sum test showed no statistically significant difference in weight between the two groups (W = 1154, p = 0.9079). Therefore, this dataset does not provide statistical evidence supporting the stereotype that snorers are heavier.


### Second Stereotype: Are snorers drinking or smoking more?

```{r}
# Compare alcohol categories between snorers and non-snorers
table(Snore$snore, Snore$alcool)

# Calculate the proportion of each alcohol category within each snoring group
alcohol_prop <- prop.table(table(Snore$snore, Snore$alcool), 1)

# Display proportions
alcohol_prop

# Visualise alcohol categories by snoring status
barplot(t(alcohol_prop),
        beside = TRUE,
        legend.text = TRUE,
        args.legend = list(title = "Alcohol category"),
        xlab = "Snoring status",
        ylab = "Proportion",
        main = "Alcohol categories by snoring status")

# Calculate mean and median alcohol values for each group
tapply(Snore$alcool, Snore$snore, mean)
tapply(Snore$alcool, Snore$snore, median)
```
```{r}
# Statistical analysis
# Check normality of alcohol values in each snoring group
shapiro.test(Snore$alcool[Snore$snore == "N"])
shapiro.test(Snore$alcool[Snore$snore == "O"])

# Compare alcohol values between the two independent groups
wilcox.test(alcool ~ snore,
            data = Snore,
            exact = FALSE)
```
**Result – Alcohol consumption**

The mean recorded alcohol value was 2.37 among non-snorers and 4.03 among snorers, while the respective medians were 0 and 4. The Wilcoxon rank-sum test revealed a statistically significant difference between the two groups (W = 788, p = 0.008594). Snorers therefore had higher recorded alcohol values in this sample.
```{r}
# Are snorers more likely to smoke?

# Compare smoking categories by snoring status
table(Snore$snore, Snore$tabac)
```

```{r}
# Proportion of smoking categories in each snoring group
smoking_prop <- prop.table(
  table(Snore$snore, Snore$tabac), 1
)

# Display proportions
smoking_prop

# Create a bar plot
barplot(t(smoking_prop),
        beside = TRUE,
        legend.text = c("Non-smokers (N)", "Smokers (O)"),
        args.legend = list(title = "Tobacco status"),
        xlab = "Snoring status",
        ylab = "Proportion",
        main = "Smoking status by snoring status")
```

```{r}
# Statistical analysis
fisher.test(table(Snore$snore, Snore$tabac))
```
**Result – Smoking habits among snorers and non-snorers**

Among non-snorers, 67.7% were smokers, compared with 57.1% among snorers.Fisher's exact test is appropriate because both variables are categorical. The test did not reveal a statistically significant association between snoring and smoking status (p = 0.3827). Therefore, the data do not provide sufficient statistical evidence to conclude that snorers are more likely to smoke.