---
title: "Testing some stereotypes"
format: pdf
Author: Tuba
---

## 2.3: Testing some stereotypes

### Loading the data

```{r}
# Import the dataset
Snore <- read.delim("snore.txt", header = TRUE)

head(Snore)
str(Snore) # structure of the dataset
summary(Snore)
```

```{r}
# Check the categories of categorical variables
unique(Snore$sex)
unique(Snore$snore)
unique(Snore$tabac)
```

```{r}
# Count observations in each category
table(Snore$sex)
table(Snore$snore)
table(Snore$tabac)
```
### First stereotype: Are snorers fatter?

```{r}
### Are snorers fatter?

# Compare weight between snorers and non-snorers
boxplot(weight ~ snore,
        data = Snore,
        xlab = "Snoring category",
        ylab = "Weight",
        main = "Weight comparison between snorers and non-snorers")
```

```{r}
# Calculate median weight in each group
tapply(Snore$weight, Snore$snore, median)
```
```{r}
# Statistical analysis
# Check normality of weight in each snoring group
shapiro.test(Snore$weight[Snore$snore == "N"])
shapiro.test(Snore$weight[Snore$snore == "O"])

# Compare weight between the two independent groups
wilcox.test(weight ~ snore,
            data = Snore,
            exact = FALSE)
```

**Result – Weight comparison** 
The median weight was 94 among non-snorers and 96 among snorers. The Wilcoxon rank-sum test showed no statistically significant difference in weight between the two groups (W = 1154, p = 0.9079). Therefore, this dataset does not provide statistical evidence supporting the stereotype that snorers are heavier.


### Second Stereotype: Are snorers drinking or smoking more?

```{r}
# Compare alcohol categories between snorers and non-snorers
table(Snore$snore, Snore$alcool)

# Calculate the proportion of each alcohol category within each snoring group
alcohol_prop <- prop.table(table(Snore$snore, Snore$alcool), 1)

# Display proportions
alcohol_prop

# Visualise alcohol categories by snoring status
barplot(t(alcohol_prop),
        beside = TRUE,
        legend.text = TRUE,
        args.legend = list(title = "Alcohol category"),
        xlab = "Snoring status",
        ylab = "Proportion",
        main = "Alcohol categories by snoring status")

# Calculate mean and median alcohol values for each group
tapply(Snore$alcool, Snore$snore, mean)
tapply(Snore$alcool, Snore$snore, median)
```
