---
title: "Testing some stereotypes"
format: pdf
Author: Tuba
---

## 2.3: Testing some stereotypes

### Loading the data

```{r}
# Import the dataset
Snore <- read.delim("snore.txt", header = TRUE)

head(Snore)
str(Snore) # structure of the dataset
summary(Snore)
```

```{r}
# Check the categories of categorical variables
unique(Snore$sex)
unique(Snore$snore)
unique(Snore$tabac)
```

```{r}
# Count observations in each category
table(Snore$sex)
table(Snore$snore)
table(Snore$tabac)
```
### First stereotype: Are snorers fatter?

```{r}
### Are snorers fatter?

# Compare weight between snorers and non-snorers
boxplot(weight ~ snore,
        data = Snore,
        xlab = "Snoring category",
        ylab = "Weight",
        main = "Weight comparison between snorers and non-snorers")
```

```{r}
# Calculate median weight in each group
tapply(Snore$weight, Snore$snore, median)
```
```{r}
# Statistical analysis
# Check normality of weight in each snoring group
shapiro.test(Snore$weight[Snore$snore == "N"])
shapiro.test(Snore$weight[Snore$snore == "O"])

# Compare weight between the two independent groups
wilcox.test(weight ~ snore,
            data = Snore,
            exact = FALSE)
```

**Result – Weight comparison** 
The median weight was 94 among non-snorers and 96 among snorers. The Wilcoxon rank-sum test showed no statistically significant difference in weight between the two groups (W = 1154, p = 0.9079). Therefore, this dataset does not provide statistical evidence supporting the stereotype that snorers are heavier.


### Second Stereotype: Are snorers drinking or smoking more?

```{r}
# Compare alcohol categories between snorers and non-snorers
table(Snore$snore, Snore$alcool)

# Calculate the proportion of each alcohol category within each snoring group
alcohol_prop <- prop.table(table(Snore$snore, Snore$alcool), 1)

# Display proportions
alcohol_prop

# Visualise alcohol categories by snoring status
barplot(t(alcohol_prop),
        beside = TRUE,
        legend.text = TRUE,
        args.legend = list(title = "Alcohol category"),
        xlab = "Snoring status",
        ylab = "Proportion",
        main = "Alcohol categories by snoring status")

# Calculate mean and median alcohol values for each group
tapply(Snore$alcool, Snore$snore, mean)
tapply(Snore$alcool, Snore$snore, median)
```
```{r}
# Statistical analysis
# Check normality of alcohol values in each snoring group
shapiro.test(Snore$alcool[Snore$snore == "N"])
shapiro.test(Snore$alcool[Snore$snore == "O"])

# Compare alcohol values between the two independent groups
wilcox.test(alcool ~ snore,
            data = Snore,
            exact = FALSE)
```
**Result – Alcohol consumption**

The mean recorded alcohol value was 2.37 among non-snorers and 4.03 among snorers, while the respective medians were 0 and 4. The Wilcoxon rank-sum test revealed a statistically significant difference between the two groups (W = 788, p = 0.008594). Snorers therefore had higher recorded alcohol values in this sample.
```{r}
# Are snorers more likely to smoke?

# Compare smoking categories by snoring status
table(Snore$snore, Snore$tabac)
```

```{r}
# Proportion of smoking categories in each snoring group
smoking_prop <- prop.table(
  table(Snore$snore, Snore$tabac), 1
)

# Display proportions
smoking_prop

# Create a bar plot
barplot(t(smoking_prop),
        beside = TRUE,
        legend.text = c("Non-smokers (N)", "Smokers (O)"),
        args.legend = list(title = "Tobacco status"),
        xlab = "Snoring status",
        ylab = "Proportion",
        main = "Smoking status by snoring status")
```

```{r}
# Statistical analysis
fisher.test(table(Snore$snore, Snore$tabac))
```
**Result – Smoking habits among snorers and non-snorers**

Among non-snorers, 67.7% were smokers, compared with 57.1% among snorers.Fisher's exact test is appropriate because both variables are categorical. The test did not reveal a statistically significant association between snoring and smoking status (p = 0.3827). Therefore, the data do not provide sufficient statistical evidence to conclude that snorers are more likely to smoke.

### Third Stereotype: Are men fatter?

```{r}
# count observations in each sex category
table(Snore$sex)
unique(Snore$sex)
```

```{r}
# create the boxplot to look at the difference of both gender's weight
boxplot(weight ~ sex,
        data = Snore,
        xlab = "Sex",
        ylab = "Weight",
        main = "Weight comparison between men and women")
```

```{r}
# calculate median weight in each sex category
tapply(Snore$weight, Snore$sex, median)

```
```{r}
# Statistical analysis

# Check normality of weight in women and men
shapiro.test(Snore$weight[Snore$sex == "F"])

shapiro.test(Snore$weight[Snore$sex == "H"])

# Compare weight between women and men
wilcox.test(weight ~ sex,
            data = Snore,
            exact = FALSE)
```

**Result – Weight comparison between men and women** The median weight was 97 among women and 94 among men. The Wilcoxon rank-sum test showed no statistically significant difference in weight between women and men (W = 893.5, p = 0.729). Thus, the data do not provide statistical evidence supporting the stereotype that men are heavier than women.

### Fourth Stereotype: Do women smoke less?

```{r}
# compare smoking status between men and women
table(Snore$sex, Snore$tabac)
```

```{r}
# calculate proportions
smoking_by_sex <- prop.table(table(Snore$sex, Snore$tabac), 1)

smoking_by_sex
```

```{r}
# create the bar plot to look at the difference of both gender's smoking status
barplot(t(smoking_by_sex),
        beside = TRUE,
        legend.text = c("Non-smokers (N)", "Smokers (O)"),
        args.legend = list(title = "Tobacco status"),
        xlab = "Sex",
        ylab = "Proportion",
        main = "Smoking status by sex")
```
```{r}
# Statistical analysis
# Test the association between sex and smoking status
fisher.test(table(Snore$sex, Snore$tabac))
```

**Result – Smoking habits among women and men**

The proportion of smokers was 40% among women and 72% among men. Fisher's exact test revealed a statistically significant association between sex and smoking status (p = 0.007276). These results indicate that women had a lower smoking prevalence than men in this sample, providing statistical support for the stereotype that women smoke less.

### General analysis: Correlations between variables

```{r}
# select numerical variables 
numerical_data <- Snore[, c("age", "weight", "height", "alcool")]
```

```{r}
# Calculate correlation matrix for selected numerical variables
cor_matrix <- cor(Snore[, c("age", "weight", "height", "alcool")],
                  use = "complete.obs")

# visualise pairwise relationships between numerical variables
pairs(Snore[, c("age", "weight", "height", "alcool")],
      main = "Relationships between numerical variables")
```
```{r}
#Statistical analysis: Spearman correlations
# Calculate the Spearman correlation matrix
spearman_matrix <- cor(numerical_data,
                       method = "spearman",
                       use = "complete.obs")

# Display Spearman correlation coefficients
spearman_matrix

# Test each pairwise Spearman correlation
correlation_pairs <- combn(names(numerical_data),
                            2,
                            simplify = FALSE)

for (pair in correlation_pairs) {
  cat("\nCorrelation between", pair[1], "and", pair[2], "\n")

  print(cor.test(numerical_data[[pair[1]]],
                 numerical_data[[pair[2]]],
                 method = "spearman",
                 exact = FALSE))
}

```

**Result – Correlation analysis** Spearman's correlation analysis revealed a very strong positive association between weight and height (rho = 0.930, p less than 0.001). This indicates that taller individuals tended to have higher body weight in this sample. No statistically significant correlations were observed between the other pairs of numerical variables (all p-values greater than 0.05). Overall, weight and height showed the strongest relationship among the variables examined.
